---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=FALSE,warning=FALSE,results='hide'}
library(dplyr,quietly=TRUE,warn.conflicts=FALSE)
library(ggplot2,quietly=TRUE,warn.conflicts=FALSE)
setwd("C:/Users/Byron/Documents/GitHub/RepData_PeerAssessment1")
Sys.setlocale("LC_TIME","English")
```
```{r}
info = read.csv('activity/activity.csv', stringsAsFactor=FALSE)
info = mutate(info, day=weekdays(as.Date(date)))
```


## What is mean total number of steps taken per day?
The fist step is to obtain the total number of steps for each day filtering the NA values:
```{r}
steps<-info%>%select(steps,date)%>%filter(!is.na(steps))%>%group_by(date)%>%summarise_each(funs(sum))
```

The histogram is obtained usign ggplot. The graphic shows that most of the days the number of steps around 10000.

```{r results='hide',warning=FALSE}
g <- ggplot(steps,aes(x=steps))
g + geom_histogram(binwidth=1000)

```

Then we can obtain the mean and the median whith the respective functions
```{r}
#Mean
mean(steps$steps)
#Median
median(steps$steps)
```


## What is the average daily activity pattern?

Plotting the number of steps in the 5 minutes interval shows the folloing graphic:

```{r}
plot(info$steps,type="l",ylab="steps/5min")

```

The maximum number of steps in a 5 minutes interval is:
```{r}
max_steps <- max(info$steps,na.rm=TRUE)
max_steps
```
This interval appears in:
```{r}
max_days <- info%>%select(steps,date,interval)%>%filter(steps==max_steps)
max_days
```

## Imputing missing values
Using the library dplyr is possible to filter the NAs, then we can count its number.
```{r}
nas = info %>% select(steps) %>% filter(is.na(steps))
nrow(nas)
```

We are going to fill replace the NAs in the data with the mean of the 5 minute steps
```{r}
mean_steps <- mean(info$steps,na.rm=TRUE)
nas = which(is.na(info$steps))
info2 <- info
info2[nas,1] <- mean_steps
```  

Once the NAs are replaced we can make a new histogram
```{r} 
steps2<-info2%>%select(steps,date)%>%group_by(date)%>%summarise_each(funs(sum))
g <- ggplot(steps2,aes(x=steps))
g + geom_histogram(binwidth=1000)
```

In this histogram we can see there are more samples around 10000 steps with respect to the last graphic.

## Are there differences in activity patterns between weekdays and weekends?

In these graphics we can se the difference in activity between the weekdays and weekends.  

The data is filtered and organized according to name of the day
```{r}
info <- mutate(info, week="test")

wkend <- which(info$day=="Saturday" | info$day=="Sunday")
info[wkend,5] <- "Weekend"

wk <- which(info$week == "test")
info[wk,5] <- "Weekday"

info_wkend <- info%>%select(steps,week)%>%filter(week=="Weekend")
info_wkend <- info_wkend%>%select(steps,week)%>%filter(!is.na(steps))
info_wkday <- info%>%select(steps,week)%>%filter(week=="Weekday")
info_wkday <- info_wkday%>%select(steps,week)%>%filter(!is.na(steps))

par(mfrow=c(2,1))
plot(info_wkend$steps,type="l",main="Weekends",ylab="steps/5 min")
plot(info_wkday$steps,type="l",main="Weekdays",ylab="steps/5 min")
```




